### 2015-06-13 writing/S'15 LSTM/intro.tex
Paper trim and proof, new figures.

Tree-structured models are presented as the more principled choice, since they directly exploit the principle of compositionality---fundamental to linguistic semantic theory \cite{Partee84,Janssen97}---which states that the meanings of phrases and sentences are constructed recursively from the meanings of their parts in a specific syntactically-guided order.

While both classes of model can perform objectively well on many tasks, and both are under active development, tree-structured models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \cite{tai2015improved} or none at all \cite{li2015tree}. 
In this paper, we use focused artificial data experiments to effectively expose a possible explanation for these underwhelming results: standard sequence models can learn to implicitly exploit recursive compositional structure.

In particular, our aim is to test standard tree and sequence models on their handling of recursive structure by training each model on sentences whose length and recursion depth are limited, and then testing each on longer and more complex sentences, such that only a model that recognizes and exploits the recursive structure will be able to succeed. An existing experimental design meets this criterion nicely: \newcite{Bowman:Potts:Manning:2014} describe an experiment and corresponding dataset which tests the ability of two types of tree-structured neural network to learn to interpret statements of a simple artificial language. We propose a way to adapt the experiment presented in that work to sequence models and use it to compare an LSTM-based model with the tree-structured models in that paper, and we extend the experiment by testing on several structured subsets of the training set in order to better understand how efficient each model is at learning the underlying generalizations.

As in \newcite{Bowman:Potts:Manning:2014}, we find that both TreeRNNs and TreeRNTNs are able to learn the necessary generalization, with their performance decaying gradually as the structures in the test set grow in size. We also find that extending the training set to include larger structures mitigates this decay. We find that a sequence model centered on a single-layer LSTM RNN is also able to generalize to unseen large structures, but that it only does this when trained on a larger and more complex training set than is needed by the tree-structured models.

@@

Recent successes on parsing by \newcite{vinyals2014grammar} and \newcite{dyer2015transition} hint that sequence models may be useful in manipulating tree structures. The simplest model presented in \newcite{vinyals2014grammar} uses an LSTM sequence model to encode each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) using only the information present in the vector, suggesting that the LSTM was able to encode the tree. However, the massive size of the training set that was necessary to train that model---250M tokens---leaves open the possibility that the model could have primarily learned to recognize and generate only tree-structures that it had already seen, representing those structures as simple hashes rather than structured objects. If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without showing demonstrating an ability to represent recursive structure, at least in a way that would generalize to substantially new structures.

>>>>

Tree-structured models are presented as the more principled choice, since they directly exploit the principle of compositionality---a fundamental idea in linguistic semantics \cite{Partee84,Janssen97}---which states that the meanings of phrases and sentences are constructed recursively from the meanings of their parts in a specific syntactically-guided order.

While both classes of model can perform objectively well on many tasks, and both are under active development, tree-structured models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \cite{tai2015improved} or none at all \cite{li2015tree}. 
In this paper, we use focused artificial data experiments to effectively expose a possible explanation for these results: standard sequence models can learn to implicitly exploit recursive compositional structure.

In particular, our aim is to test standard tree and sequence models on their handling of recursive structure by training each model on sentences whose length and recursion depth are limited, and then testing each on longer and more complex sentences, such that only models that recognize and exploit the recursive structure will be able to succeed. An existing experimental design meets this criterion nicely: \newcite{Bowman:Potts:Manning:2014} describe an experiment and corresponding dataset which tests the ability of two tree-structured models to learn to interpret statements of a simple artificial language. We propose a way to adapt than experiment to sequence models and use it to compare an LSTM-based model with the three tree-structured models and we extend the experiment by testing on several structured subsets of the training set in order to better understand how efficiently each model learns the underlying generalizations.

As in \newcite{Bowman:Potts:Manning:2014}, we find that standard tree-structured neural networks are able to learn the necessary generalization, with their performance decaying gradually as the structures in the test set grow in size. We also find that extending the training set to include larger structures mitigates this decay. In addition, we find that a sequence model centered on a single-layer LSTM is also able to generalize to unseen large structures, but that it only does this when trained on a larger and more complex training set than is needed by the tree-structured models.

@@

Recent successes on parsing by \newcite{vinyals2014grammar} and \newcite{dyer2015transition} hint that sequence models may be able to exploit recursive structure. The simplest model presented in Vinyals et al.~uses an LSTM sequence model to encode each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) using only the information present in the vector, suggesting that the LSTM was able to encode the tree. However, the massive size of the training set that was necessary to train that model---250M tokens---leaves open the possibility that the model primarily learned to recognize and generate only tree-structures that it had already seen, representing those structures as simple hashes rather than structured objects. If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without showing demonstrating an ability to represent recursive structure, at least in a way that would generalize to substantially new structures.
